---
title: "Spatial Growth Workup"
author: "Maia Kapur"
format: html
editor: source
toc: true
toc-depth: 2
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(growthbreaks)
library(dplyr)
library(ggplot2)
codspec <- read.csv(here::here('data','survey_specimen_data_pcod.csv'))
names(codspec) <- tolower(names(codspec))
# names(codspec)[12] <- 'length_cm'
## split by sex
codspec <- codspec %>% 
  filter(sex != 3) %>%
  dplyr::select(sex, year, age, length, lat = end_latitude , long = end_longitude, weight) %>%
  group_split(sex) 
theme_set(ggsidekick::theme_sleek())
```

# Overview

This analysis leverages the [growthbreaks](https://github.com/afsc-assessments/growthbreaks) R package to investigate potential spatio-temporal breaks, or clusters, in Pacific cod growth across the range. There are some major caveats in analyzing these data wholesale, largely due to sampling changes through time -- so this investigation is intended to guide discussions and inference, *not* to be the final word on spatial demographic structure. This uses `SPECIMEN` data from all three regions. 

 

# Raw Observations
Males = 1, Females = 2
It looks like ages 3-9 are good candidates for break detection.
```{r echo = FALSE, include = TRUE, message = FALSE, warning = FALSE}
dat_use <- data.frame(codspec[[1]])
pm <- check_data(dat_use, showPlot = FALSE)

ggplot(bind_rows(codspec[[1]],codspec[[2]]),aes(x = age, y =length) ) +
  geom_point(alpha = 0.1) +
  facet_grid(~sex)

 
```
 

```{r female breaks, warning = FALSE, message=FALSE, echo = FALSE, include = FALSE}
breakpointsF <- get_Breaks(dat = data.frame(codspec[[2]]),
                          ages_to_use = c(3,5,7,9),
                          sex = FALSE, 
                          axes = 0, 
                          showPlot = FALSE)
## the longitude break is meaningless, delete a priori
breakpointsF$breakpoints$long <- -Inf
breakpointsF$breakpoints <- breakpointsF$breakpoints[2:3,]
breakpointsF$breakpoints$lat[1] <- 53

fits1f <- refit_Growth( data.frame(codspec[[2]]), breakpointsF$breakpoints, showPlot = FALSE)

plot_Breaks(dat =  data.frame(codspec[[2]]),
            breakpoints = breakpointsF$breakpoints,
            TRUE)

```

```{r male breaks, warning = FALSE, message=FALSE, echo = FALSE, include = FALSE}
breakpointsM <- get_Breaks(dat = data.frame(codspec[[1]]),
                          ages_to_use = c(3,5,7,9),
                          sex = FALSE, 
                          axes = 0, 
                          showPlot = TRUE)

## the longitude break is meaningless, delete a priori
breakpointsM$breakpoints$long<- -Inf
fits1m <- refit_Growth( data.frame(codspec[[1]]), breakpointsM$breakpoints, showPlot = FALSE)

fits1m$fits_plot
fits1m$pars_plot


plot_Breaks(dat =  data.frame(codspec[[1]]),
            breakpoints = breakpointsM$breakpoints,
            TRUE)
```


::: panel-tabset

## Male Detected Breaks
```{r, echo = FALSE, message = FALSE, warning = FALSE}
plot_Breaks(dat =  data.frame(codspec[[1]]),
            breakpoints = breakpointsM$breakpoints,
            TRUE)
```

## Male fitted curves

```{r, echo = FALSE, message = FALSE, warning = FALSE}
fits1m$fits_plot

```

## Male parameters
```{r, echo = FALSE, message = FALSE, warning = FALSE}
fits1m$pars_plot
```



## Female Detected Breaks
```{r, echo = FALSE, message = FALSE, warning = FALSE}
plot_Breaks(dat =  data.frame(codspec[[2]]),
            breakpoints = breakpointsF$breakpoints,
            TRUE)
```

## Female fitted curves

```{r, echo = FALSE, message = FALSE, warning = FALSE}
fits1f$fits_plot

```

## Female parameters
```{r, echo = FALSE, message = FALSE, warning = FALSE}
fits1f$pars_plot
```




:::
